\textbf{ Matching the forecast horizon with the relevant ecological processes }

### 2019-07-22 doc/manuscript.tex
Emphasize biological processes in Intro

Ecological models used for forecasting have traditionally been based on either time-series modeling or space-for-time substitutions.
Time-series approaches use the historical behavior of a system to fit models that describe the temporal dynamics of that system and use those
dynamic models to make predictions about what will happen in the future. This is a common approach for population based forecasts and includes both
process based population models and data-driven models capturing empirical patterns in data (e.g., \citep{ward_complexity_2014}) 
Alternatively, space-for-time modeling approaches develop models of how and ecological property varies in across space in response to environmental
conditions. These spatial relationships are assumed to also hold for changes at a site through time, and forecasts are made by forecasting future
environmental conditions and determining the associated ecological response. Most forecasts for future species distributions and biodiversity
are based on this type of approach, with a models built to describe the environmental conditions where species occur across space being applied to
forecast environmental conditions to determine where they will occur in the future (e.g., citation). Despite both of these approaches being
widely used there has been little comparison of how they compare for forecasting and understanding the dynamics of ecological systems \citep{harris_forecasting_2018}.

Whether historical dynamics, contemporary spatial patterns, or a combination of the two will serve as the best source of information for forecasting
may depend on how far into the future we are attempting to forecast \citep{harris_forecasting_2018}. This potential dependency on the "forecast horizon"; ()\textit{sensu} citation)
results from lags in the response of ecological conditions to environmental change, changes in the importance of ecological processes with
time scale \citep{levin_1992,rosenzweig_1995}, and differences between time-series and spatial gradients in the range of conditions present in the data.
At short forecast horizons (days to years) conditions are likely to stay within the range of the historical data at a site, the time scales of change will be similar to
the time scales present in that historical data, and the current state of the system is likely to capture the influence of unmeasured processes.
As a result, for near-term forecasts time-series approaches may effectively capture the key dynamics and result in accurate forecasts. In contrast,
at long forecast horizons (decades) an ecosystem is likely to experience conditions that have not been historically observed () and be responding to
processes operating at scales difficult to detect in even long ecological time-series. Therefore, for the kinds of far-term forecasts typically conducted
in biodiversity and species-distribution modeling space-for-time based approaches may include important information about how ecosystems respond to major
shifts in climate over long periods and as a results produce better long-term forecasts. Using different approaches for different time horizons is common
in other areas of forecasting, including forecasts for environmental conditions where models for short-term changes (weather) differ substantially from
those for long-term changes (climate).

>>>>

Ecological models used for forecasting have traditionally been based on either time-series modeling or space-for-time substitutions.
Time-series approaches use the historical behavior of a system to fit models that describe the temporal dynamics of that system and use those
dynamic models to make predictions about what will happen in the future. This is a common approach for population based forecasts and includes both
process based population models and data-driven models capturing empirical patterns in data (e.g., \citep{ward_complexity_2014}). Time-series models
capture processes like birth, death, small scale dispersal events, and short-term responses to environmental conditions.
Alternatively, space-for-time modeling approaches develop models of how and ecological property varies in across space in response to environmental
conditions. These spatial relationships are assumed to also hold for changes at a site through time, and forecasts are made by forecasting future
environmental conditions and determining the associated ecological response. Most forecasts for future species distributions and biodiversity
are based on this type of approach, with a models built to describe the environmental conditions where species occur across space being applied to
forecast environmental conditions to determine where they will occur in the future (e.g., citation). Space-for-time models capture processes like
immigration, extinction, and responses to large changes in environmental conditions over long time periods. Despite both of these approaches being
widely used there has been little comparison of how they compare for forecasting and understanding the dynamics of ecological systems \citep{harris_forecasting_2018}.

Whether historical dynamics, contemporary spatial patterns, or a combination of the two will serve as the best source of information for forecasting
may depend on how far into the future we are attempting to forecast \citep{harris_forecasting_2018}. This potential dependency on the "forecast horizon"; ()\textit{sensu} citation)
results from lags in the response of ecological conditions to environmental change, changes in the importance of ecological processes with
time scale \citep{levin_1992,rosenzweig_1995}, and differences between time-series and spatial gradients in the range of conditions present in the data.
At short forecast horizons (days to years) processes driving individual demographics tend to dominate (similar to the time scales present in that
historical data), conditions are likely to stay within the range of the historical data at a site, and the current state of the system is likely
to capture the influence of unmeasured processes. As a result, for near-term forecasts time-series approaches may effectively capture the key dynamics
and result in accurate forecasts. In contrast, at long forecast horizons (decades) an ecosystem is likely to experience conditions that have not been
historically observed () and be responding to large scale processes large scale climate and land use change that are difficult to detect in even long
ecological time-series. Therefore, for the kinds of long-term forecasts typically conducted in biodiversity and species-distribution modeling
space-for-time based approaches may include important information about how ecosystems respond to major shifts in climate over long periods and as a
results produce better long-term forecasts. Using different approaches for different time horizons is common in other areas of forecasting,
including forecasts for environmental conditions where models for short-term changes (weather) differ substantially from those for long-term changes
(climate).
